Go Implementation of Up-To Techniques for Equivalence of Weighted Languages

### 2020-09-30 paper/paper.tex
fixed bug in svd nullspace computation
annihilator, go code cleanup

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage[table]{xcolor}
\usepackage{stmaryrd}
\usepackage{amsfonts}

@@

\begin{defn}
  \textbf{} \\

@@

The algorithm is defined by the iterative method:
\begin{eqnarray}
  R_0 = \mker{o}^0, & \quad & R_{i+1} = R_i + \sum_{a \in A} t_a(R_i)^t \label{back} 
\end{eqnarray}
The algorithm stops when $R_{j+1} = R_j$. Proof is available in section 4.2 of \cite{BONCHI201277}

@@

The algorithms and data structures for this paper are implemented in the Go programming 
language. This implementation makes use of the \textit{Gonum} library for numerical 
computations. We only import the Gonum libraries for matrices and linear algebra 
and visual plotting of samples and functions.

Real numbers are implemented with double precision floating point numbers,  
known as the \texttt{float64} type in the Go programming language.

@@

\begin{defn}
  \textbf{Algorithm for computing the nullspace} \\
  The algorithm implementation can be found in file \texttt{lin/nullspace.go}; it is
  adapted from \cite{scipy/ranknullspace}. 
  Let's consider the singular value decomposition of a matrix $A \in \R^{m \times n}$:

@@

\begin{exmp}
  The actual Go implementation is quite verbose, we can show a shorter Python 
  implementation of the algorithm to compute the nullspace, using the 
  \textit{SciPy} library \cite{scipy/ranknullspace}:

  \begin{lstlisting}[language=python]
def nullspace(A, atol=1e-13, rtol=0):
    A = np.atleast_2d(A)
    u, s, vh = svd(A)
    tol = max(atol, rtol * s[0])
    nnz = (s >= tol).sum()
    ns = vh[nnz:].conj().T
    return ns
    \end{lstlisting}
\end{exmp}

>>>>

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage[table]{xcolor}
\usepackage{stmaryrd}
\usepackage{amsfonts}
\usepackage{lstlistings-golang}

@@

% TODO linear weighted automata
\begin{defn}
  \textbf{} \\

@@

The algorithm is defined by the iterative method:
\begin{eqnarray}
  R_0 = \mker{o}^0, & \quad & R_{i+1} = R_i + \sum_{a \in A} t_a(R_i)^t \label{back} 
\end{eqnarray}
Where $\mker{o}^0$ is an annihilator.
The algorithm stops when $R_{j+1} = R_j$. Proof is available in section 4.2 of \cite{BONCHI201277}

@@

The algorithms and data structures for this paper are implemented in the Go programming 
language. This implementation makes use of the \textit{Gonum} library for numerical 
computations. We only import the Gonum libraries for matrices and linear algebra 
and visual plotting of samples and functions.
Real numbers are implemented with double precision floating point numbers,  
known as the \texttt{float64} type in the Go programming language.

@@

\begin{defn}
  \textbf{Algorithm for computing the null space of a vector subspace} \\
  The algorithm implementation can be found in file \texttt{lin/nullspace.go}; it is
  adapted from \cite{scipy/ranknullspace}. 
  Let's consider the singular value decomposition of a matrix $A \in \R^{m \times n}$:

@@

\begin{exmp}
  First, we show a shorter Python 
  implementation of the algorithm to compute the nullspace, using the 
  \textit{SciPy} library \cite{scipy/ranknullspace}:

  \lstinputlisting[language=python]{nullspace.py}

    The Go implementation is quite longer:

    \lstinputlisting[language=go]{../lin/nullspace.go}

\end{exmp}

\subsection{Implementing the backwards algorithm}

In \cite{BONCHI201277}, for the first step of the iterative backwards 
algorithm to compute the largest linear weighted bisimulation, we 
need to compute $\mker{o}^0$. If $V$ is a vector space and $W$ is a
subspace of $W$, the annihilator of $W$, respectively $W^0$ is 
a subspace of the space $V^*$ of linear functionals on $V$.
$W^0$ are the functionals that annihilate on $W$. Since 
we are working on subspaces of $\R^n$, we can directly compute 
the orthogonal complement in our implementation instead of the
annihilator.

\begin{prop}
  Within finite dimensional spaces, the annihilator $W^0$ is isomorphic
  to the orthogonal complement.  
\end{prop}

\begin{proof}
  Let $V$ be an inner product space over the field $\K$ with an inner product
  $\langle \cdot , \cdot \rangle : V \times V \to \K$ defined. 
  With the inner product defined on $V$, every linear functional can be 
  represented with a vector: if $\xi : V \to \K$ is a functional, $c$ 
  is an unique vector such that $\forall x \in V \: . \: \langle c, x \rangle = \xi(x)$.
  The map associating every $\xi \in V^*$ with the vector $c \in V$ that corresponds 
  to $\xi$ with respect to the inner product is a linear isomorphism. Through this 
  isomorphism, the annihilator $W^0$ becomes $W^\perp$.
\end{proof}

To compute the orthogonal complement of a vector subspace $W$, we
compute $W^\perp = \mker{A^T}$, where $A$ is the matrix with 
column vectors in the spanning set of $W$ as its columns. Precisely, $W$ is 
represented as the 
column space of $A$. Proof is available in \cite{ila}.
